---
title: "Analysis of Red Wine Quality "
author: "Mohamed Alsayed"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document: default
  html_document:
    # theme of html document
    # theme of code highlight                                 
    # table of contents
    theme       : cosmo            # "default", "cerulean", "journal",
                                    # "flatly", "readable", "spacelab",
                                    # "united", "cosmo", "lumen", "paper", 
                                    # "sandstone", "simplex", "yeti"
    highlight   : tango          # "default", "tango", "pygments",
                                    # "kate",  "monochrome", "espresso",
                                    # "zenburn", "haddock", "textmate"
    toc         : true              # get table of content
    toc_depth   : 3
    toc_float   : true
    code_folding: hide
    number_sections: yes
    fig_caption: yes
    df_print    : paged
---

______________

```{r echo=FALSE, message=FALSE, warning=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

library(ggplot2)
library(knitr)
library(dplyr)

# knitr: Suppress code/messages/warnings 
#        Set default plot options and center them
knitr::opts_chunk$set(fig.width=9,fig.height=5,fig.path='Figs/',
                      fig.align='center',tidy=TRUE,
                      echo=FALSE,warning=FALSE,message=FALSE)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
red_wine <- read.csv('WineQualityReds.csv')
```


 This tidy data set contains 1,599 red wines with 11 variables on the chemical
properties of the wine. At least 3 wine experts rated the quality of each wine,
providing a rating between 0 (very bad) and 10 (very excellent).

# Univariate Plots Section


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
summary(red_wine$quality)

# Histogram - Quality
ggplot(aes(x=quality), data= red_wine) + 
  geom_histogram(stat="count") +
  scale_y_continuous(breaks=seq(0, 700, 100)) +
  scale_x_continuous(breaks=seq(3, 8, 1))
```


The quality histogram shows that the majority of wine samples are rated 5 or 6 
while few samples are rated 1 and 8 and no sample rated above 8 nor less than 3.



```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(red_wine$fixed.acidity)

# Histogram - fixed.acidity
ggplot(aes(x=fixed.acidity), data= red_wine) + 
  geom_histogram(binwidth = 0.1) +
  scale_x_continuous(breaks=seq(4, 16, 1))

```

 The fixed acidity histogram show that the majority of wine samples has fixed acidity of apporximately 7 to 8 and the rest ranges between 4.5 to 13 and few of the samples has higher fixed acidity above 13 to 16.


```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(red_wine$volatile.acidity)

# Histogram - volatile.acidity
ggplot(aes(x=volatile.acidity), data= red_wine) + 
  geom_histogram(binwidth = 0.01) +
  scale_x_continuous(breaks=seq(0.2, 1.6, 0.1))
```

 The volatile acidity histogram show that the majority of wine samples has volatile acidity of 0.3 to 0.7


```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(red_wine$citric.acid)

# Histogram - citric.acid
ggplot(aes(x=citric.acid), data= red_wine) + 
  geom_histogram(binwidth = 0.01)
```

The Citric Acid histogram shows that the majority of wine samples ranges
between 0 and 0.75 and also we can observe high number of these samples has 0
citric acid as well as 0.5 citric acid.


```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(red_wine$residual.sugar)

# Histogram - residual.sugar
ggplot(aes(x=residual.sugar), data= red_wine) + 
  geom_histogram(binwidth = 0.1) +
  scale_x_continuous(breaks=seq(1, 16, 1))
  
```


The Residual Sugar histogram shows that most of wine samples are between 2
and 3 on residual sugar and some few samples above 4 to 16.



```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(red_wine$chlorides)

# Histogram - chlorides
ggplot(aes(x=chlorides), data= red_wine) + 
  geom_histogram(binwidth = 0.01) +
  scale_x_continuous(breaks=seq(0.0, 0.6, 0.05))
```

The Chlorides histogram shows that the majority of wine samples are between 
0.05 to 0.1 on the chlorides scale.



```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(red_wine$free.sulfur.dioxide)

# Histogram - free.sulfur.dioxide
ggplot(aes(x=free.sulfur.dioxide), data= red_wine) + 
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks=seq(0, 70, 10))
```

The Free Sulfur Dioxide histogram shows that the wine samples ranges 
between 0 to 50 and many samples has 5 value on the free sulfur dioxide scale.


```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(red_wine$total.sulfur.dioxide)

# Histogram - total.sulfur.dioxide
ggplot(aes(x=total.sulfur.dioxide), data= red_wine) + 
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks=seq(0, 300, 20))
```
The total Sulfur Dioxide histogram shows that the majority of wine samples ranges 
between 10 to 120 and there are some outliers above 260 .


```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(red_wine$density)

# Histogram - density
ggplot(aes(x=density), data= red_wine) + 
  geom_histogram(bins=30) +
  scale_x_continuous(breaks=seq(0.99, 1.10, 0.0010))
```

The density histogram shows a nearly normal distrbuted data with a peak of 0.997 and
range between 0.990 to 1.004


```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(red_wine$pH)

# Histogram - pH
ggplot(aes(x=pH), data= red_wine) + 
  geom_histogram(binwidth = 0.01) +
  scale_x_continuous(breaks=seq(2, 4, 0.1))
```

The Ph histogram shows that most wine sample has ph value that ranges 
between 3 to 3.7 and some of the outlier samples are above 4.


```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(red_wine$sulphates)

# Histogram - sulphates
ggplot(aes(x=sulphates), data= red_wine) + 
  geom_histogram(stat="count")
```

 The sulphates histogram shows that most of wine samples are ranging between 
0.5 to 1.


```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(red_wine$alcohol)

# Histogram - alcohol
ggplot(aes(x=alcohol), data= red_wine) + 
  geom_histogram(binwidth = 0.1) +
  scale_x_continuous(breaks=seq(9, 14, 1))
```


The Alcohol histogram shows that the wine samples has alcohol ranges between
9 to 14 with a high number of samples contains 9.5 of alcohol.



# Univariate Analysis


### What is the structure of your dataset?
The dataset consist of 11 variables that contribute in the red wine quality like
alcohol, ph and density.

### What is/are the main feature(s) of interest in your dataset?
The most interesting variable is alcohol since its the main factor 
contributing directly to the quality of the red wine.


### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
The ph value as well as the density of the wine will both be good factors to 
detrmine the wine quality.


### Did you create any new variables from existing variables in the dataset?
I measured every variable independently to be able to see its effect on the 
quality and no new variables has been added.


### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
The data is pretty tidy and has no errors nor nulls but distributions just
needed some adjustment for readability so I changed the bin width for most of
histograms and adjusted their axis for better observations. 


# Bivariate Plots Section



```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}

# jitter and boxplot showing relation between quality and alcohol 
ggplot(aes(x=as.factor(quality), y=alcohol), data= red_wine) + 
  geom_jitter(alpha= 0.5 , width = 0.35, height = 0) +
  geom_boxplot(color ='blue', alpha=0.3)
```

 The plot above shows the relation between alcohol and the quality of red
wine, and we can observe that the quality increases when alcohol increases.


```{r echo=FALSE, message=FALSE, warning=FALSE}

# jitter and boxplot showing relation between quality and pH 
ggplot(aes(x=as.factor(quality), y=pH), data= red_wine) + 
  geom_jitter(alpha= 0.5 , width = 0.35, height = 0) +
  geom_boxplot(color ='blue', alpha=0.3)
```

The relation between pH and the quality are week and the plot shows
that high pH can get nearly same result as lower pH in quality.

```{r echo=FALSE, message=FALSE, warning=FALSE}

# jitter and boxplot showing relation between quality and density 
ggplot(aes(x=as.factor(quality), y=density), data= red_wine) + 
  geom_jitter(alpha= 0.5 , width = 0.35, height = 0) +
  geom_boxplot(color ='blue', alpha=0.3)

```

The plot above showing the relation between quality and density, we can observe
that there is an inverse relation when density increases the quality decreases
and vice versa.



# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

-Alcholol has big impact on the quality level of the red wine, we can see from 
the plotted data that the increase of alchohol increases the chance of high
quality rates.
-Density has inverse relation with the quality when the density increases 
the quality decreases
-The pH value of the red wine did not show much relation and effect on the 
quality rate of the red wine.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Yes, both density and pH values and how both impacted the change in quality
rates across all red wine samples.


### What was the strongest relationship you found?
-Alcohol has strongest relationship to the quality rate of the red wine.


# Multivariate Plots Section


```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}

# colored point plot showing relation between quality, alcohol and density
ggplot(aes(x=alcohol, y=density, color= as.factor(quality)), data= red_wine) + 
  geom_point(alpha= 0.5 , width = 0.7, height= 0)  +
  scale_color_brewer(palette = "RdYlBu", name="Quality",
                     direction=-1)

```

From the scatterplot above we can observe the inverse relation between density
and alcohol, when density increases the alcohol decreases. we can also see that
the quality of red wine is assosicated with high level of alcohol and lower 
density values which proves previous investigation.

```{r echo=FALSE, message=FALSE, warning=FALSE}


# colored point plot showing relation between quality, alcohol and pH
ggplot(aes(x=alcohol, y=pH, color= as.factor(quality)), data= red_wine) + 
  geom_point(alpha= 0.5 , width = 0.7, height= 0)  +
  scale_color_brewer(palette = "RdYlBu", name="Quality",
                     direction=-1)

```


The above scatterplot shows a weak relationship between pH value and alcohol
most samples scores between 3 to 3.5 pH and associated with 10 alcohol value
and the increase in alcohol associated with more quality rates.



# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

The Density, alcohol and quality scatterplot strenghtened the previous observed
data from the bivariate section and showed how both of those two factors
contributed in the red wine quality rating.

### Were there any interesting or surprising interactions between features?
The pH value of the red wine has no effect on the quality results nor even
has a clear relationship with alcohol which is surprising.



------

# Final Plots and Summary


### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}

# Histogram - quality
ggplot(aes(x=as.factor(quality)), data= red_wine) + 
  geom_histogram(stat="count", fill= 'blue') +
  scale_y_continuous(breaks=seq(0, 700, 100)) +
  scale_x_discrete(breaks=seq(3, 8, 1)) +
  labs(x='Quality (score between 0 and 10)', y='Counts' , title='Quality of red wine Samples')

```

### Description One
The quality histogram shows that the majority of wine samples are rated 5 or 6 
while few samples are rated 1 and 8 and no sample rated above 8 nor less than 3.



### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}

# jitter and boxplot showing the relation between quality and alcohol
ggplot(aes(x=as.factor(quality), y=alcohol), data= red_wine) + 
  geom_jitter(alpha= 0.5 , width = 0.3 , height= 0) + geom_boxplot(color='blue',alpha=0.3) +
  labs(x='Quality (score between 0 and 10)', y='Alcohol (% by volume)', 
       title='relation between Alcohol and Quality of Red wine') +
  scale_y_continuous(breaks=seq(8, 16, 1)) +
  scale_x_discrete(breaks=seq(2, 9, 0.5)) 
```

### Description Two
 The plot above shows the relation between alcohol and the quality of red
wine, and we can observe that the quality increases when alcohol increases.

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}

# colored point plot showing relation between quality, alcohol and density
ggplot(aes(x=alcohol, y=density, color= as.factor(quality)), data= red_wine) + 
  geom_point(alpha= 0.5 , width = 1.5) +
  labs(x='Alcohol (% by volume)', y='expression(Density~(g/cm^{3}))', 
       title= 'Relation between Density, Alcohol and Quality of Red wine' ) +
  scale_y_continuous(breaks=seq(0.990, 1.1, 0.005)) +
  scale_x_continuous(breaks=seq(8, 16, 1)) +
  scale_color_brewer(palette = "RdYlBu", name="Quality", direction=-1)
```

### Description Three

The above scatterplot shows a weak relationship between pH value and alcohol
most samples scores between 3 to 3.5 pH and associated with 10 alcohol value
and the increase in alcohol associated with more quality rates.

------

# Reflection

 The Red wine database contains around 1600 sample of red wine data and all of
their chemical combinations, with these data we performed a high level data 
analysis to understand how much every variable of these chemicals contribute
the most in each sample quality.

Our observation from these data that some variables has big impact on the
quality rating and some other not for instance, alcohol showed a propotional
relationship toward the quality whenever the alcohol increases the quality 
increases. Another factor which has big impact on quality is the density of
the red wine sample it has inverse relationship with the quality, when the 
density increases the quality decreases and the alcohol value as well.

For the pH value is surprising since it does not have any impact on the quality
our plots shows no matter high or less the pH value the quality remain the
same.

Our analysis is limited to the small data set we observed our data from as
well as the information we have for each sample as well as using lineear
regression model only.

Furthermore, for future analysis we might use logestic regression analysis
for better analysing the red wine quality.